Generalized Bradley-Terry models

### 2025-02-20 inst/jss_hyper3_real.Rnw
more work including using data(sushi)

We now apply the ideas above to a dataset obtained from the {\tt
  preflib} library.

<<defcalculate_sushi_H}>>=
make_sushi_H <- function(a, lambda){

    sushi_types <- c("ebi", "anago", "maguro", "ika", "uni", "sake", "tamago", "toro", "tekka", "kappa")
    ec <- # equivalence classes
        c(`ebi`  = 1, `anago` = 2, `maguro` = 1, `ika` = 1, `uni` = 1,
          `sake` = 1, `tamago` = 1, `toro` = 2, `tekka-maki` = 2,
          `kappa-maki` = 1)

    H <- hyper3()

    for(i in seq_len(nrow(a))){
        H <- H + cheering3(sushi_types[a[i,]], e=ec, h=c(1, lambda))
    }
    return(H)
}
@

<<readsushi>>=
a <- as.matrix(read.table("sushi.txt",skip=22))
wanted <- sample(seq_len(nrow(a)), 5, replace=TRUE, prob=a[,1])
a <- a[wanted, -1]
@

<<tryH,cache=TRUE>>=

H1888 <- make_sushi_H(a, strength=1.888)
H1000 <- make_sushi_H(a, strength=1.000)
@

>>>>

We now apply the ideas above to a dataset obtained from the {\tt
  preflib} library~\citep{MaWa13a} in which ten different types of
sushi were placed in order of preference by 5000 judges.  Three of the
sushis (maguro, toro, tekka-maki) were made with tuna, and it is
reasonable to view these three types as being exchangeable in much the
same way as the red and blue buses.  To proceed, we need to define
equivalence classes of sushi---one for tuna-based sushis and one for
everything else.  Of course, one can also imagine the opposite
phenomenon whereby the judge decides he has had enough tuna for the
moment and this disincentivises further tuna choices.  This mechanism
would be modelled by having $\lambda<1$.

<<defcalculate_sushi_H>>=

ec <- c("ebi"  = 1, "anago" = 2, "maguro" = 1, "ika" = 1, "uni" = 1,
        "sake" = 1, "tamago" = 1, "toro" = 2, "tekka-maki" = 2,
        "kappa-maki" = 1)

make_sushi_H <- function(lambda){    
    for(i in seq_len(nrow(a))){
        H <- H + cheering3(sushi_types[a[i,]], e=ec, h=c(1, lambda))
    }
    return(H)
}
@
